### 2022-06-22 manuscript/manuscript.Rmd
further updates to manuscript
- gams + effect derivatives in fine and gross motor domains

@@

When modelling the effect of chronological age corrected for prematurity on developmental domains, we observed a significant non-linear relationship on communication scores ($F_{smooth}$ (5.2, 224.04) = 13.43, *p* < 0.001), that reflect an overall negative marginal effect ($\beta$ = -2.36, CI~95%~[-3.47, -1.25], $t_{student}$ (224.04) = -4.2, *p* < 0.001), however, this was not true when assessing the direction of the effect in the age range between 1 to 7 ($\beta$ = 0.49, CI~95%~[-0.89, 1.86], $t_{student}$ (224.04) = 0.45, *p* = 0.319), and neither in the 18 to 48 months old group ($\beta$ = 0.45, CI~95%~[-1.32, 2.23], $t_{student}$ (224.04) = 0.42, *p* = 0.593), whereas the effect tend to be positive but non-significant.

@@

When analysing the motor skills domain, we found a significant non-linear effect of corrected age on gross motor scores ($F_{smooth}$ (5.24, 226.75) = 6.19, *p* < 0.001), which had an overall positive effect ($\beta$ = 1. 95, CI~95%~[0.66, 3.25], $t_{student}$ (226.75) = 2.97, *p* = 0.003), however, the slope varied as a function of age, with a negative effect on gross motor score $\beta$ = -2.94, CI~95%~[-4.55, -1. 34], $t_{student}$$ (226.75) = -3.7, *p* = 0.004, but in the range 9.7 to 15.5 a positive effect was observed ($\beta$ = 1.86, CI~95%~[0.61, 3.11], $t_{student}$ (226.75) = 2.93, *p* = 0.009), while in the range 7.3 to 9.2 ($\beta$ = 0.02, CI~95%~[-1. 12, 1.17]) and 16 to 48 months ($\beta$ = -0.02, CI~95%~[-2.05, 2.01]) the slope was non-significant and virtually zero ($t_{student}$ (226.75) = 0.06, *p* = 0.45, and $t_{student}$ (226.75) = 0.07, *p* = 0.646 respectively).

>>>>

## Communication

@@

When modelling the effect of chronological age corrected for prematurity on developmental domains, we observed a significant non-linear relationship on communication scores ($F_{smooth}$ (5.2, 224.04) = 13.43, *p* < 0.001), that reflect an overall negative marginal effect ($\beta$ = -2.36, CI~95%~[-3.47, -1.25], $t_{student}$ (224.04) = -4.2, *p* < 0.001), however, this was not true when assessing the direction of the effect in the age range between 0 to 6.8 ($\beta$ = 0.49, CI~95%~[-0.89, 1.86], $t_{student}$ (224.04) = 0.45, *p* = 0.319), and neither in the 18.4 to 48 months old group ($\beta$ = 0.45, CI~95%~[-1.32, 2.23], $t_{student}$ (224.04) = 0.42, *p* = 0.593), whereas the effect tend to be positive but non-significant.

## Motor skills

@@

When analysing the motor skills domain, we found a significant non-linear effect of corrected age on gross motor scores, $F_{smooth}$ (5.24, 226.75) = 6.19, *p* < 0.001, which had an overall positive effect ($\beta$ = 1.95, CI~95%~[0.66, 3.25], $t_{student}$ (226.75) = 2.97, *p* = 0.003), however, the slope varied as a function of age, with a negative effect in the 0 to 6.8 age range ($\beta$ = -2.94, CI~95%~[-4.55, -1.34], $t_{student}$ (226.75) = -3.7, *p* = 0.004), but in the 9.7 to 15.5 interval, this relationship was inverted ($\beta$ = 1.86, CI~95%~[0.61, 3.11], $t_{student}$ (226.75) = 2.93, *p* = 0.009), although, in the 7.3 to 9.2 ($\beta$ = 0.02, CI~95%~[-1.12, 1.17]) and 16 to 48 age range ($\beta$ = -0.02, CI~95%~[-2.05, 2.01]) the slope was non-significant and virtually zero ($t_{student}$ (226.75) = 0.06, *p* = 0.45, and $t_{student}$ (226.75) = 0.07, *p* = 0.646 respectively).

```{r eval=FALSE}
mod <- gam(
  motora_fina_total ~ 
    s(profesional_id, sexo_paciente, respondedor_vinculo, bs = "re") + 
    s(edad_corregida_meses), 
  data = dataset, 
  method = "REML"
)

report_overall(mod, "edad_corregida_meses")
#> [1] "$\\beta$ = 0.04, CI~95%~[-0.45, 0.52], $t_{student}$ (226.77) = 0.14, *p* = 0.886"

report_slopes(mod, "edad_corregida_meses")
#>    Start  End edad_corregida_meses                                                                                 expr
#> 1:   0.0 21.8                 10.9    $\\beta$ = 0.03, CI~95%~[-0.54, 0.61], $t_{student}$ (226.77) = 0.04, *p* = 0.434
#> 2:  22.3 38.3                 30.3 $\\beta$ = -0.79, CI~95%~[-1.45, -0.12], $t_{student}$ (226.77) = -2.34, *p* = 0.022
#> 3:  38.8 48.0                 43.4  $\\beta$ = -0.88, CI~95%~[-1.97, 0.21], $t_{student}$ (226.77) = -1.62, *p* = 0.113

report_smooth(mod)
#>                                                   term                                           expr
#> 1:                             s(edad_corregida_meses) $F_{smooth}$ (2.59, 226.77) = 4.2, *p* = 0.005
#> 2: s(profesional_id,sexo_paciente,respondedor_vinculo) $F_{smooth}$ (3.63, 226.77) = 0.7, *p* = 0.021
```

The same was observed when inspecting the effect of corrected age in the fine motor domain scores ($F_{smooth}$ (2.59, 226.77) = 4.2, *p* = 0.005), however it was not possible to estimate a significant overall effect ($\beta$ = 0.04, CI~95%~[-0.45, 0.52], $t_{student}$ (226.77) = 0.14, *p* = 0.886), nevertheless, it was only in the 22.3 to 38.3 age range where a significant negative effect was observed ($\beta$ = -0.79, CI~95%~[-1.45, -0.12], $t_{student}$ (226.77) = -2.34, *p* = 0.022).
